Disturbance Assimilation

### 2020-11-19 Paper.Rmd
small text update, pulled out Discussion to own file

# Discussion

* Interpretation
  * TBD based on results

* Implications  (borrow text from CMS proposals)
  * Improving MRV
  * Recovery forecasting
  * Disturbance nowcasting
    * requires reducing temporal resolution

* Future directions
  * Disturbance priors need to be spatially heterogenous and evolve temporally
    * Connect to state-and-transition literature
  * Scaling up -> spatially implicit
    * ED2
    * Contagion

>>>>

<!---- DISCUSSION --->
```{r child = 'discussion.Rmd'}
```

### 2020-11-19 methods.Rmd
small text update, pulled out Discussion to own file

@@

3. Simulated data experiments (known disturbance)
  + Single Site, Single Disturbance
  + SSSD Multi Pool
  + SS Multi-Disturbance w/ PFT switching
  + Simple Multi-site: spatial cov
4. Real-world examples
  + Describe Model & PEcAn
  + Calibrate state-and-transition prior
    + Landtrendr
    ftp://islay.ceoas.oregonstate.edu/change_agent_attribution
  + Applications: PNW
    + Logging
    + Fire
    + Land use transition (forest -> non?)

>>>>

*Priors*

  $$X^L_k \sim MVN(\vec{\mu}_{f,k},P_{f,k})$$
  $$\vec{\rho} \sim Multinomial(1,\vec{p})$$ 
*Process Model*

  $$X = X^L \vec{\rho}$$

*Likelihood*
  $$Y \sim MVN(X,R)$$
### Ensemble Adjustment

[@Anderson2001]

*reassignment of classes*

To reassign forecasted ensemble members to new disturbance classes we first identify those that decreased between the prior probability, $p_k$, and the posterior frequency, $\bar{\rho}_k$.

$$\delta p_k = max\{0,\bar{\rho}_k-p_k \}$$

Classes that decreased in relative abundance are reassigned to new classes according to a Multinomial draw

$$c_{A|k} = Multinomial\left( n_k, {{\delta p}\over{\sum \delta p}}  \right)$$

*Ensemble adjustment by class*

@@

3. Simulated data experiments (known disturbance)
  + x Single Site, Single Disturbance  (add a case where EnKF has disturb but it's more rare)
  + x SSSD Multi Pool
  + x SS Multi-Disturbance - distinct
  + x SS Multi-Disturbance w/ distinct, PFT switching
      * mention parallel PFT ensembles
      * need to protect against long-term PFT leakage; related to more general need to "look back"
  + Simple Multi-site: spatial cov
  + SSMD - initially overlapping but diverge
      * how to handle updating of assignment based on new information?
  + SSMD - transition probabilities change with state
  + rate of false pos/neg: affected by p.dist, R, Q, P0 -- how much do we need to simulate?
4. Real-world examples
  + Describe Model & PEcAn
  + Calibrate state-and-transition prior
    + Landtrendr
    ftp://islay.ceoas.oregonstate.edu/change_agent_attribution
  + Applications: PNW
    + Logging
    + Fire
    + Land use transition (forest -> non?)
